---
title: "Exercises for class 1"
output: html_document
date: "13-02-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rethinking)
```
```{r}
install.packages(c("coda","mvtnorm","devtools","loo","dagitty","shape"))
devtools::install_github("rmcelreath/rethinking")
```



# Exercises for Class 1

## Supporting code

Below is code from Chapter 2 of the book that you can use to do the exercises. Not all the code you require is here though. Make sure to not rush it and take time to play around with each function. :))

```{r}
# Grid Approximation

# define grid
p_grid = seq(from = 0, to = 1, length.out=20)

# define prior
prior = rep(1, 20)

# compute likelihood at each value in grid
likelihood = dbinom(6, size = 9, prob = p_grid)

# compute product of likelihood and prior
unstd.posterior = likelihood * prior

# standardize the posterior so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

# plot the distribution
plot(p_grid, posterior, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")

```

## Exercises

### Easy

2E1. 

```{r}

(2)

```

2E2.

```{r}

(3)

```

2E3.

```{r}

(1)

```

2E4.


De Finetti's Subjective Probability

Bruno de Finetti was a strong proponent of subjective Bayesian probability.  This means that he didn't see probabilities as something inherent to objects or events in the world. Instead, he believed probability represents a degree of belief given a particular state of knowledge. Different people, with different information, can have different probabilities for the same event.

The Globe Tossing Example

When we say, "the probability of the globe landing on water is 0.7," a de Finettian interpretation is:

Not an objective fact: It's not saying that the globe itself has some intrinsic property that makes it land on water 70% of the time.
Betting behavior: Saying the probability is 0.7 means you'd be willing to bet on water with certain odds. For instance, you might bet $7 to win $3 if the globe lands on land. This reflects your degree of belief based on your available knowledge (maybe the Earth's surface composition).
Subject to change: If we're presented with new information (e.g., someone secretly added more water to the globe), our belief, and therefore the probability we assign, could change.
The Importance of the Observer

In the globe tossing example, the following things highlight why probability is subjective in de Finetti's view:

Knowledge of the globe: If someone unfamiliar with Earth estimates the probability of water, they might guess 0.5 (50/50) due to lacking specific information.
Bias: If someone strongly believes most of the planet is land, their subjective probability of water might be lower, even if they know the actual proportions.
Key Takeaway

De Finetti's perspective  emphasizes that probability isn't about the nature of a "random" event like tossing a globe.  Instead, it reflects the way we reason about the world given incomplete information. This makes probability a powerful tool for making decisions under uncertainty, but it's  crucial to remember its subjective roots.



### Medium

2M1.


### First senario 
```{r}
# Write your code here.
# define grid
p_grid <- seq(from = 0, to = 1, length.out = 20)

# define uniform prior
prior <- rep(1, 20)

# compute likelihood for W, W, W
likelihood <- dbinom(3, size = 3, prob = p_grid)

# compute unstandardized posterior
unstd.posterior <- likelihood * prior

# standardize the posterior
posterior <- unstd.posterior / sum(unstd.posterior)

# plot
plot(p_grid, posterior, type = "b", xlab = "probability of water", ylab = "posterior probability",
     main = "Posterior for W, W, W")


```
### Second scenario

```{r}
# compute likelihood for W, W, W, L
likelihood <- dbinom(3, size = 4, prob = p_grid)

# compute unstandardized posterior
unstd.posterior <- likelihood * prior

# standardize the posterior
posterior <- unstd.posterior / sum(unstd.posterior)

# plot
plot(p_grid, posterior, type = "b", xlab = "probability of water", ylab = "posterior probability",
     main = "Posterior for W, W, W, L")

```
```{r}
# compute likelihood for L, W, W, L, W, W, W
likelihood <- dbinom(5, size = 7, prob = p_grid)

# compute unstandardized posterior
unstd.posterior <- likelihood * prior

# standardize the posterior
posterior <- unstd.posterior / sum(unstd.posterior)

# plot
plot(p_grid, posterior, type = "b", xlab = "probability of water", ylab = "posterior probability",
     main = "Posterior for L, W, W, L, W, W, W")

```

2M2.

```{r}
# define grid
p_grid <- seq(from = 0, to = 1, length.out = 20)

# define step-like prior
prior <- ifelse(p_grid < 0.5, 0, 1)

# Scenario 1: Observations W, W, W
likelihood1 <- dbinom(3, size = 3, prob = p_grid)
unstd.posterior1 <- likelihood1 * prior
posterior1 <- unstd.posterior1 / sum(unstd.posterior1)
plot(p_grid, posterior1, type = "b", xlab = "probability of water", ylab = "posterior probability",
     main = "Posterior for W, W, W with step-like prior")

# Scenario 2: Observations W, W, W, L
likelihood2 <- dbinom(3, size = 4, prob = p_grid)
unstd.posterior2 <- likelihood2 * prior
posterior2 <- unstd.posterior2 / sum(unstd.posterior2)
plot(p_grid, posterior2, type = "b", xlab = "probability of water", ylab = "posterior probability",
     main = "Posterior for W, W, W, L with step-like prior")

# Scenario 3: Observations L, W, W, L, W, W, W
likelihood3 <- dbinom(5, size = 7, prob = p_grid)
unstd.posterior3 <- likelihood3 * prior
posterior3 <- unstd.posterior3 / sum(unstd.posterior3)
plot(p_grid, posterior3, type = "b", xlab = "probability of water", ylab = "posterior probability",
     main = "Posterior for L, W, W, L, W, W, W with step-like prior")

```

2M3.

Suppose there are two globes, one for Earth and one for Mars. The Earth globe is 70% covered
in water. The Mars globe is 100% land. Further suppose that one of these globes—you don’t know
which—was tossed in the air and produced a “land” observation. Assume that each globe was equally
likely to be tossed. Show that the posterior probability that the globe was the Earth, conditional on
seeing “land” (Pr(Earth|land)), is 0.23.

```{r}

pr_earth = 0.5

pr_mars = 0.5

pr_land_given_earth = 0.3

pr_land_given_mars = 1


# Calculate the total probability of observing land
pr_land = pr_earth * pr_land_given_earth + pr_mars * pr_land_given_mars

# Calculate the posterior probability of the Earth given land
pr_earth_given_land = (pr_land_given_earth * pr_earth) / pr_land

print(pr_earth_given_land)

```
The posterior probability that the globe was the Earth, conditional on seeing "land" (Pr(Earth∣land)), is approximately 0.23.



For the exercises below, I highly suggest you to grab a piece of paper and try to solve the problems in a 'visual' way.'


Suppose you have a deck with only three cards. Each card has two sides, and each side is either
black or white. One card has two black sides. The second card has one black and one white side. The
third card has two white sides. Now suppose all three cards are placed in a bag and shuffled. Someone
reaches into the bag and pulls out a card and places it flat on a table. A black side is shown facing up,
but you don’t know the color of the side facing down. Show that the probability that the other side is
also black is 2/3. Use the counting method (Section 2 of the chapter) to approach this problem. This
means counting up the ways that each card could produce the observed data (a black side facing up
on the table).




2M4.

```{r}
# Define the scenarios
# BB = Black/Black, BW = Black/White
# We only consider these two because WW (White/White) cannot show a black side up

# The possible outcomes where a black side is facing up
possible_outcomes <- c('BB1', 'BB2', 'BW')

# Count the total number of outcomes where a black side is up
total_black_up <- length(possible_outcomes)

# Count the outcomes where the other side is also black
# Only the BB card can have the other side black, and it has 2 sides that can be up
black_on_both_sides <- sum(substr(possible_outcomes, 1, 2) == 'BB')

# Calculate the probability
probability_black_other_side <- black_on_both_sides / total_black_up

# Print the probability
print(paste("Probability that the other side is also black:", probability_black_other_side))

```

2M5.Now suppose there are four cards: B/B, B/W, W/W, and another B/B. Again suppose a card is
drawn from the bag and a black side appears face up. Again calculate the probability that the other
side is black.

```{r}

# Define the scenarios considering the additional BB card
# BB1, BB2 are the two Black/Black cards, and BW is the Black/White card
possible_outcomes <- c('BB1_1', 'BB1_2', 'BB2_1', 'BB2_2', 'BW')

# Count the total number of outcomes where a black side is up
total_black_up <- length(possible_outcomes)

# Count the outcomes where the other side is also black
# Both BB cards can have the other side black, each with 2 sides that can be up
black_on_both_sides <- sum(substr(possible_outcomes, 1, 2) == 'BB')

# Calculate the probability
probability_black_other_side <- black_on_both_sides / total_black_up

# Print the probability
print(paste("Probability that the other side is also black:", probability_black_other_side))

```

2M6. Imagine that black ink is heavy, and so cards with black sides are heavier than cards with white
sides. As a result, it’s less likely that a card with black sides is pulled from the bag. So again assume
there are three cards: B/B, B/W, and W/W. After experimenting a number of times, you conclude that
for every way to pull the B/B card from the bag, there are 2 ways to pull the B/W card and 3 ways to
pull the W/W card. Again suppose that a card is pulled and a black side appears face up. Show that
the probability the other side is black is now 0.5. Use the counting method, as before.

```{r}
# Define the weighted chances for pulling each card
weighted_chances <- c(BB = 1, BW = 2, WW = 3)

# Calculate weighted outcomes for a black side being up
weighted_outcomes_black_side_up <- weighted_chances['BB'] * 2 + weighted_chances['BW'] * 1

# Calculate weighted outcomes where the other side is also black (only possible for BB)
weighted_outcomes_black_both_sides <- weighted_chances['BB'] * 2

# Calculate the probability
probability_black_other_side <- weighted_outcomes_black_both_sides / weighted_outcomes_black_side_up

# Print the probability
print(paste("Probability that the other side is also black:", probability_black_other_side))


```
